---
title: "Stock Price Predictions Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
runtime: shiny
---

```{r global, include=FALSE}

# Function to Install and Load R Packages
install.load.packages <- function(required.packages)
{
  required.packages <-
    required.packages[!(required.packages %in% installed.packages()[, "Package"])]
  
  
  if (length(required.packages))
  {
    install.packages(required.packages, repos = 'http://cran.us.r-project.org');
  }
  for (package.name in required.packages)
  {
    library(package.name,
            character.only = TRUE,
            quietly = TRUE)
    
  }
}

# Specify the list of required packages to be installed and load
required.packages <- c(
  "flexdashboard",
  "quantmod",
  "rpart",
  "nlme",
  "xts",
  "DMwR",
  "TTR",
  "nnet",
  "randomForest",
  "gridExtra",
  "openxlsx",
  "bizdays",
  "shiny",
  "DT"
)


# Call the Function
install.load.packages(required.packages)

library(flexdashboard)
library(quantmod) 
library(rpart) 
library(nlme)
library(xts)
library(DMwR)
library(TTR)
library(nnet)
library(randomForest)
library(gridExtra)
library(openxlsx)
library(shiny)
library(DT)

```

```{r, include=FALSE}

# Aux functions used to obtain predictors
myATR   <- function(x) ATR(HLC(x))[,'atr']
myADX   <- function(x) ADX(HLC(x))[,'ADX']
myEMV   <- function(x) EMV(cbind(Hi(x),Lo(x)),Vo(x))[,2]
myMACD  <- function(x) MACD(Cl(x))[,2]
mySAR   <- function(x) SAR(cbind(Hi(x),Cl(x))) [,1]
myVolat <- function(x) volatility(OHLC(x),calc="garman")[,1]

assign("myATR", myATR, envir = .GlobalEnv)
assign("myADX", myADX, envir = .GlobalEnv)
assign("myEMV", myEMV, envir = .GlobalEnv)
assign("myMACD", myMACD, envir = .GlobalEnv)
assign("mySAR", mySAR, envir = .GlobalEnv)
assign("myVolat", myVolat, envir = .GlobalEnv)

#Obtain prediction model for a given iteration
GetPredModel <- function(i, stockName){
  formula <- as.formula(paste0("Next(Cl(",stockName,")",",",i,") ~
                                Delt(Cl(",stockName,"),k=1)+
                                myATR(",stockName,")+
                                myADX(",stockName,")+
                                myEMV(",stockName,")+
                                myVolat(",stockName,")+
                                myMACD(",stockName,")+
                                mySAR(",stockName,")+
                                runMean(Cl(",stockName,"))"))
  m <- specifyModel(formula)
  data <- as.data.frame(modelData(m))
  return(data)
} 

```



Column {.sidebar}
-----------------------------------------------------------------------

```{r}

# Select dropdowns in the side bar
selectInput('ticker', 'Select a Stock', c("Pfizer - PFE", 
                                          "Goldman Sachs - GS", 
                                          "General Motors - GM", 
                                          "Apple - AAPL", 
                                          "Amazon - AMZN"))

selectInput('modelID', 'Select a Predictive Model', c("ANN", 
                                                      "RandomForest",
                                                      "Rpart"))

```


row {data-height=650}
-----------------------------------------------------------------------

### Stock Price & Volume Movement Chart (past 180 days) 

```{r}
# Select ticker based on user selection 
stock.ticker <- reactive({
  if (input$ticker == "Goldman Sachs - GS") {
    ticker = "GS"
  } else if (input$ticker == "General Motors - GM") {
    ticker = "GM"
  } else if (input$ticker == "Pfizer - PFE") {
    ticker = "PFE"
  } else if (input$ticker == "Apple - AAPL") {
    ticker = "AAPL"
  } else {
    ticker = "AMZN"
  }
  ticker
})
```


```{r}

# Render candle chart based on the ticker

renderPlot({
  
  stockName = stock.ticker()
  modelID = input$modelID
  
  # end date and start dates based on system time
  endDate = as.Date(Sys.Date() - 1, format = "%Y-%M-%D")
  startDate = as.Date(endDate - 180, format = "%Y-%M-%D")
  
  # Plot the actual stock price
  chartData <-
    getSymbols(
      stockName,
      src = "yahoo",
      from = startDate,
      to = endDate,
      auto.assign = FALSE
    )
  chartSeries(
    chartData,
    name = input$ticker,
    subset = paste(endDate - 180, endDate, sep = "/"),
    type = 'line',
    theme = 'white'
  )
  
})

```

Row {data-height=450}
-------------------------------------

### Stock Price Predictions (next 5 trading days) based on choice of predictive model 

```{r}

# Calculate forecast and plot forecast
renderPlot({
  
  stockName = stock.ticker()
  modelID = input$modelID
  
  if (input$modelID == "Rpart") {
    modelID = 1
  } else if (input$modelID == "RandomForest") {
    modelID = 2
  } else {
    modelID = 3
  }
  
  # End date and start dates based on system time
  endDate = as.Date(Sys.Date() - 1, format = "%Y-%M-%D")
  startDate = as.Date(endDate - 180, format = "%Y-%M-%D")
  
  # Derive max data available
  maxdate <-
    max(index(getSymbols(stockName, auto.assign = FALSE)))
  
  # Define prediction as xts object
  prediction <- xts()
  n = 1
  while (n <= 5) {
    data <- GetPredModel(n, stockName)
    formula <-
      as.formula(paste0("Next.Cl.", stockName, ".", n, " ~ ."))
    set.seed(1234)
    
    # Pick model
    if (modelID == 1) {
      res <- rpart(formula, data, parms = list(prior = c(.65,.35), split = "information"))
    } else if (modelID == 2) {
      res <- randomForest(formula, data, na.action = na.omit, ntree = 50)
    } else if (modelID == 3) {
      res <-
        nnet(
          formula,
          data,
          size = 10,
          decay = 0.01,
          maxit = 1000,
          linout = T,
          trace = F
        )
    }
    
    pred_date <- as.Date(maxdate) + n
    p <- predict(res, data[as.character(maxdate),])
    if (modelID == 3) {
      p <- p[, 1]
    }
    
    #print(pred_date)
    names(p) <- c(as.character(pred_date))
    prediction <- rbind(as.xts(cbind(p)), prediction)
    n = n + 1
  }
  data.preds <- reactive({
    as.data.frame(data)
  }) 
  plot(prediction)
  
})

```

### Summary Table | Best Performing Models

QuantMod package has been used extensively for getting the tickers, specifying the model and building the model. Evaluated models based on performance estimations for each stock. Based on the average theil metric, we have predicted the closing stock price for each ticker for the next 5 days. Below is a table explaining the best model for each metric and their average MSE, MAE and THIEL for the selected stock ticker based on the top performance metrics:

```{r}
# Show table with metrics in the summary section 
renderTable({
  t.data <- read.xlsx("./ticker_data.xlsx")
  t.data <- as.data.frame(t.data)
  
  if (input$ticker == "Goldman Sachs - GS") {
    df <- t.data[1,3:8]
  } else if (input$ticker == "General Motors - GM") {
    df <- t.data[2,3:8]
  } else if (input$ticker == "Pfizer - PFE") {
    df <- t.data[3,3:8]
  } else if (input$ticker == "Apple - AAPL") {
    df <- t.data[4,3:8]
  } else {
    df <- t.data[5,3:8]
  }
  df
  
})

```

